---
title: "CH4 analyses"
author: "Abby Lewis"
date: "2026-01-20"
output: html_document
---

## Overview

This file contains code to analyze and visualize the effects of soil warming on methane emissions, including changes in the proportional contribution of extreme fluxes (i.e., top 5%, 1%, and 0.2%). Code chunks should be run in order.

### Table of contents: 

-   Step 1: Load data and packages
-   Step 2: Calculate summary statistics
-   Step 3: Plot data summary (Figure 2 and variants)
-   Step 4: Analyze the contribution of extreme fluxes (Figure 3 and variants)
-   Step 5: Simulate low-frequency sampling to assess bias (Figure 4)
-   Step 6: Daytime and WL sampling biases (SI)
-   Step 7: Analyze residuals for annual and extreme fluxes


## Step 1: Load data and packages

```{r setup}
# Load packages and functions
library(tidyverse)
source("R/hm_style.R")
source("R/plot_all_data.R")
source("R/make_gam_label.R")
source("R/make_lm_label.R")
source("R/calc_bias.R")

# Load data
ch4 <- read_csv(here::here("01_Data", "GENX_flux_data_interpolated_with_drivers.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels),
         date = as_date(DateTime_EST)
  )

soil_temp <- read_csv(here::here("01_Data", "Soil_temp.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels))

smartx <- read_csv(here::here("01_Data", "SMARTX CH4 and CO2 fluxes_2016-2025.csv")) %>%
  mutate(date = as_date(Date, format = "%m/%d/%Y"),
         CH4_umol_m2_h = `CH4 flux (umol/m2/d)`/24)

# Number of days in analysis
n_days <- as.numeric(difftime(max(ch4$DateTime_EST),
  min(ch4$DateTime_EST),
  units = "days"
))

annual_totals <- read_csv("annual_totals.csv")
```
```{r}
# Only use days that have data from all 12 chambers
hm_df <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12)

# Calculate the contribution of extreme fluxes
annual_totals <- hm_df %>%
  mutate(
    month = month(date),
    season = recode_values(month,
                           3:5 ~ "Spring",
                           6:8 ~ "Summer",
                           9:11 ~ "Autumn")) %>%
  group_by(Chamber, season) %>%
  mutate(n_days = as.numeric(difftime(max(DateTime_EST),
                                      min(DateTime_EST), units = "days"))) %>%
  # Calculate sum of fluxes that exceed each threshold
  # and convert to growing season total (mol/m2/y from umol/m2/h)
  summarize(
    n = sum(!is.na(CH4_umol_m2_h)),
    annual_flux = sum(CH4_umol_m2_h, na.rm = T) /
      n / 1000 / 1000 * (unique(n_days) * 24),
    n_days = unique(n_days)
  )
```


## Step 5: Simulate low-frequency sampling to assess bias (Figure 4)

```{r fig_4}
# Set simulation specs
reps <- 10000
set.seed(4700)
intervals <- c("daily", "twice weekly", "weekly", "fortnightly", "monthly")

# Add labels for day, month, week, and fortnight
df <- ch4 %>%
  mutate(
    day = yday(date),
    month = month(date),
    week = week(date),
    fortnight = week %/% 2
  ) %>%
  mutate(season = recode_values(month,
                                3:5 ~ "Spring",
                                6:8 ~ "Summer",
                                9:11 ~ "Autumn"))

# Run simulations
df_out <- purrr::map(intervals, calc_bias, season = T) %>% # See calc_bias.R
  bind_rows()

# Summarize outputs
to_plot <- df_out %>%
  group_by(Chamber, rep, interval, time, season) %>%
  summarize(
    mean_hourly = sum(flux * n / sum(n), na.rm = TRUE),
    tot_hourly_not_yet_mult = mean_hourly / 1e6, # Growing season total (simulated)
    .groups = "drop"
  ) %>%
  left_join(
    annual_totals %>% rename(tot_real = annual_flux), # Add totals (calculated high-frequency)
    by = c("Chamber", "season")
  ) %>%
  mutate(tot_hourly = tot_hourly_not_yet_mult * n_days * 24, # Growing season total (simulated)
         percent_bias = tot_hourly / tot_real * 100 - 100,
         # Add temperature treatment
         Temp_C = sub(" ºC", "", 
                      str_extract(as.character(Chamber), 
                                  "-*[0-9]+\\.*[0-9]* ºC")),
         Temp_C = as.numeric(sub("mb", "0", Temp_C))
  )

to_plot_factored <- to_plot %>%
  mutate(season = factor(season,
                         levels = c("Spring", "Summer", "Autumn")))

# Calculate medians
to_plot_median <- to_plot_factored %>%
  group_by(Chamber, interval, time, Temp_C, season) %>%
  summarize(
    median_bias = median(percent_bias, na.rm = TRUE),
    .groups = "drop"
  )

# Plot medians by time interval
sims_full_day <- to_plot_median %>%
  filter(time == "all") %>%
  mutate(
    interval = factor(interval, levels = intervals),
    int_num = as.numeric(interval)
  ) %>%
  ggplot(aes(color = Chamber, x = int_num, y = median_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_line(size = 0.5) +
  geom_point(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5) +
  scale_x_continuous(
    breaks = 1:length(intervals),
    labels = intervals
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  facet_wrap(~season) +
  ylab("Manual sampling bias (%)") +
  xlab("Simulated sampling interval") +
  guides(fill = guide_legend(ncol = 2)) +
  theme(axis.text.x = element_text(
    hjust = 1,
    angle = 45
  ),
  legend.title.position = "top")

# Plot monthly medians by temperature
# First, calculate statistics for plot
sims_label <- make_gam_label(
  data = to_plot_median %>%
    filter(
      interval == "monthly",
      time == "all"
    ),
  xvar = "Temp_C",
  yvar = "median_bias",
  group_var = "season"
)

# Generate plot
sims_month <- to_plot_median %>%
  filter(
    interval == "monthly",
    time == "all"
  ) %>%
  ggplot(aes(x = as.numeric(Temp_C), y = median_bias, group = "group")) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML"),
              color = "black") +
  geom_point(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5) +
  geom_text(
    data = sims_label,
    aes(x = -Inf, y = -Inf, label = label),
    hjust = 0, vjust = -0.5,
    inherit.aes = FALSE, size = 2.5
  ) +
  scale_x_continuous(
    limits = c(-0.5, 6.5),
    expand = c(0, 0)
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  facet_wrap(~season) +
  theme(legend.title.position = "top") +
  ylab("Manual sampling bias (%)") +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 2))

# Plot full distribution
sims_dist <- to_plot_factored %>%
  filter(
    interval == "monthly",
    time == "all"
  ) %>%
  ggplot(aes(x = as.numeric(Temp_C), y = percent_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  # Without quantiles for shading
  geom_violin(aes(color = Chamber, fill = Chamber),
    position = "identity", alpha = 0.1, width = 0.5
  ) +
  # With 25 and 75%
  geom_violin(aes(color = Chamber), fill = NA,
    quantiles = c(0.25, 0.75), quantile.linetype = "11",
    position = "identity", width = 0.5
  ) +
  # With median
  geom_violin(aes(color = Chamber), fill = NA,
    quantiles = c(0.5), quantile.linetype = "solid",
    position = "identity", width = 0.5
  ) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  theme_hm +
  facet_wrap(~season)+
  scale_x_continuous(
    limits = c(-0.5, 6.5),
    expand = c(0, 0)
  ) +
  theme(legend.title.position = "top") +
  ylab("Manual sampling bias (%)") +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 2))

# Combine plots
with_dist2 <- patchwork::wrap_plots(
  a = sims_month +
    guides(fill = guide_legend(ncol = 1)) +
    ylab("Monthly sampling bias\n(median %)") +
    xlab("Temperature treatment (ºC above ambient)"),
  b = sims_dist + theme(legend.position = "none") +
    ylab("Monthly sampling bias (%)"),
  c = sims_full_day +
    ylab("Sampling bias\n(median %)") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(hjust = 1, angle = 45)
    ),
  d = patchwork::guide_area(),
  guides = "collect",
  axes = "collect_x",
  design = c("AAABBB
              AAABBB
              CCCCDD
              CCCCDD")
) +
  patchwork::plot_annotation(
    tag_levels = "a"
  ) &
  theme(plot.tag = element_text(face = "bold"))

# Visualize and save
with_dist2
ggsave(with_dist2,
  filename = "../03_Figures/SI/Figure 4a - Simulations all by season.jpg",
  dpi = 300, width = 8, height = 5.5
)
```


## Step 6: Daytime and WL sampling biases (SI)

```{r fig_4}
# Set simulation specs
intervals <- c("monthly")

# Add labels for day, month, week, and fortnight
df <- ch4 %>%
  mutate(
    day = yday(date),
    month = month(date),
    week = week(date),
    fortnight = week %/% 2,
    DateTime_local = with_tz(DateTime_EST, tzone = "America/New_York"))%>%
  mutate(season = recode_values(month,
                                3:5 ~ "Spring",
                                6:8 ~ "Summer",
                                9:11 ~ "Autumn"))

# Run simulations
df_out <- calc_bias("monthly", only_time = F, season = T) %>% # See calc_bias.R
  bind_rows()

# Summarize outputs
to_plot <- df_out %>%
  group_by(Chamber, rep, interval, time, season) %>%
  summarize(
    mean_hourly = sum(flux * n / sum(n), na.rm = TRUE),
    tot_hourly_not_mult = mean_hourly / 1e6, # Growing season total (simulated)
    .groups = "drop"
  ) %>%
  left_join(
    annual_totals %>% rename(tot_real = annual_flux), # Add totals (calculated high-frequency)
    by = c("Chamber", "season")
  ) %>%
  mutate(tot_hourly = tot_hourly_not_mult * (n_days * 24),
         percent_bias = tot_hourly / tot_real * 100 - 100,
         # Add temperature treatment
         Temp_C = sub(" ºC", "", 
                      str_extract(as.character(Chamber), 
                                  "-*[0-9]+\\.*[0-9]* ºC")),
         Temp_C = as.numeric(sub("mb", "0", Temp_C)),
         season = factor(season,
                         levels = c("Spring", "Summer", "Autumn")))

# Calculate medians
to_plot_median <- to_plot %>%
  group_by(Chamber, interval, time, Temp_C, season) %>%
  summarize(
    median_bias = median(percent_bias, na.rm = TRUE),
    .groups = "drop"
  )

# Stats
to_plot_median %>%
  group_by(interval, time, season) %>%
  summarize(median = median(median_bias))

# Plot medians by time interval
sims_multiple_biases <- to_plot_median %>%
  mutate(
    time = factor(time, 
                  levels = c("all", "daytime", "waterlevel"),
                  labels = c("Fully random\nsampling times", 
                             "Only daytime\nsampling", 
                             "Only sampling\nwhen water level\nis at or below the\nmarsh surface")),
  ) %>%
  ggplot(aes(x = time, y = median_bias)) +
  geom_hline(yintercept = 0, color = "grey85") +
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(aes(fill = Chamber), shape = 21, size = 2, stroke = 0.5, width = 0.2) +
  scale_fill_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  scale_color_manual(
    values = color.gradient,
    name = "Chamber"
  ) +
  coord_cartesian(xlim = c(0.5, 3.5), expand = c(left = F, right = F))+
  theme_hm +
  facet_wrap(~season)+
  ylab("Bias of monthly sampling (%)") +
  xlab("Simulated sampling bias") +
  guides(fill = guide_legend(ncol = 6)) +
  theme(axis.title.x = element_blank(),
        legend.box.margin = margin(t = -25),
        legend.title.position = "top")

sims_multiple_biases

ggsave(sims_multiple_biases,
  filename = "../03_Figures/SI/SI - Multiple biases by season.jpg",
  dpi = 300, width = 9, height = 4
)
```