---
title: "Other stats"
author: "Abby Lewis"
date: "2024-08-22"
output: html_document
---

## Overview

This file contains code to calculate relevant statistics for manuscript text.

### Table of contents:

-   Step 1: Load data and packages
-   Step 2: Calculate inundation and salinity


## Step 1: Load data and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(tidyverse)
library(patchwork)
library(mgcv)

# Load data
ch4 <- read_csv(here::here("01_Data", "GENX_flux_data_interpolated_with_drivers.csv")) %>%
  mutate(DateTime_EST = with_tz(DateTime_EST, tz = "EST"),
         Chamber = factor(Chamber, levels = chamber_labels),
         date = as_date(DateTime_EST)
  )

n_days <- as.numeric(difftime(max(ch4$DateTime_EST),
  min(ch4$DateTime_EST),
  units = "days"
))
```

## Step 2: Calculate ebullition

```{r}
ch4 %>%
  group_by(Chamber) %>%
  summarize(n_eb = sum(Ebullition_yn, na.rm = T),
            pct_eb = n_eb / sum(!is.na(Ebullition_yn)),
            n_tot = sum(!is.na(Ebullition_yn)))
```

## Step 4: Analyze the contribution of extreme fluxes (Figure 3)

```{r fig_3}
# Only use days that have data from all 12 chambers
hm_df <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12)

# Calculate the contribution of extreme fluxes
hot_moments_data <- hm_df %>%
  group_by(Chamber) %>%
  # Identify whether fluxes exceed each threshold
  mutate(
    hot_moment_quant_998 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.998, na.rm = TRUE
      ),
    "hot", "not"
    ),
    hot_moment_quant_99 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.99, na.rm = TRUE
      ),
    "hot", "not"
    ),
    hot_moment_quant_95 = ifelse(CH4_umol_m2_h >=
      quantile(CH4_umol_m2_h,
        probs = 0.95, na.rm = TRUE
      ),
    "hot", "not"
    ),
    hot_moment_eb = ifelse(Ebullition_yn,
    "hot", "not"
    )
  ) 

hot_moments <- hot_moments_data %>%
  # Calculate sum of fluxes that exceed each threshold
  # and convert to growing season total (mol/m2/y from umol/m2/h)
  summarize(
    n = sum(!is.na(CH4_umol_m2_h)),
    annual_flux = sum(CH4_umol_m2_h, na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_998 = sum(CH4_umol_m2_h[hot_moment_quant_998 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_99 = sum(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_95 = sum(CH4_umol_m2_h[hot_moment_quant_95 == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    hot_sum_eb = sum(CH4_umol_m2_h[hot_moment_eb == "hot"], na.rm = T) /
      n / 1000 / 1000 * (n_days * 24),
    min = min(DateTime_EST),
    max = max(DateTime_EST)
  ) %>%
  mutate( # Add temperature treatment
    Temp_C = sub(" ºC", "", str_extract(as.character(Chamber), "-*[0-9]+\\.*[0-9]* ºC")),
    Temp_C = as.numeric(sub("mb", "0", Temp_C))
  )

# Plot overall growing season totals
annual <- hot_moments %>%
  ggplot(aes(x = Temp_C, y = annual_flux)) +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML"),
              color = "black") +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Growing season", "emissions (mol" ~ m^-2 ~ yr^-1 * ")"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  theme(
    legend.position = "bottom",
    legend.title.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 1))

# Calculate summary stats about the contribution of extreme events
top_pct <- hot_moments %>%
  mutate(
    hm_pct_998 = hot_sum_998 / annual_flux * 100,
    hm_pct_99 = hot_sum_99 / annual_flux * 100,
    hm_pct_95 = hot_sum_95 / annual_flux * 100,
    hm_pct_eb = hot_sum_eb / annual_flux * 100
  ) %>%
  select(Temp_C, Chamber, hm_pct_998, hm_pct_99, hm_pct_95, hm_pct_eb) %>%
  pivot_longer(c(hm_pct_95, hm_pct_99, hm_pct_998, hm_pct_eb)) %>%
  mutate(
    name = factor(
      name,
      levels = c("hm_pct_95", "hm_pct_99", "hm_pct_998", "hm_pct_eb"),
      labels = c(
        "Top 5% of measurements",
        "Top 1% of measurements",
        "Top 0.2% of measurements",
        "Ebullition events"
      )
    )
  ) %>%
  group_by(name) %>%
  mutate(
    model = list(
      mgcv::gam(
        value ~ s(Temp_C, k = 3),
        data = cur_data(),
        method = "REML"
      )
    ),
    p      = summary(model[[1]])$s.pv,
    r2     = summary(model[[1]])$r.sq,
    median = median(value)
  ) %>%
  ungroup()

# Format summary stats for plot
label_df <- top_pct %>%
  group_by(name) %>%
  summarise(
    p = first(p),
    r2 = first(r2),
    median = first(median),
    label = paste0(
      "  Median = ", round(median, 1), " %",
      "\n  R² = ", round(r2, 2),
      "\n  p = ", round(p, 3)
    )
  )

# Plot the contribution of extreme events to total emissions
top_pct_fig <- top_pct %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x, k = 3),
    method.args = list(method = "REML"),
    color = "black"
  ) +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  geom_text(
    data = label_df,
    aes(x = -Inf, y = Inf, label = label),
    hjust = 0, vjust = 1.5,
    inherit.aes = FALSE, size = 3
  ) +
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Percent of growing season", "emissions"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name, scales = "free_y")

# Combine growing season sums with extreme events
comb <- ggpubr::ggarrange(annual + 
                            theme(legend.position = "right"),
  top_pct_fig + theme(legend.position = "none"),
  nrow = 2, ncol = 1, heights = c(1, 1.5),
  labels = "auto"
)

# Save
ggsave(
  comb,
  filename = "../03_Figures/Manuscript/Figure 3??? - Top percent eb.jpg",
  dpi = 300, width = 5, height = 6
)
```

Ebullition figure

```{r}
names(color.gradient) <- chamber_labels
chamber_scale <- scale_fill_manual(
  limits = chamber_labels,
  values = color.gradient,
  name = "Chamber",
  drop = FALSE
)

eb_number <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12) %>%
  group_by(Chamber) %>%
  summarize(
    n = sum(Ebullition_yn, na.rm = TRUE),
    pct = n / sum(!is.na(Ebullition_yn)) * 100
  ) %>%
  mutate(
    Temp_C = sub(
      " ºC", "",
      str_extract(
        as.character(Chamber),
        "-*[0-9]+\\.*[0-9]* ºC"
      )
    ),
    Temp_C = as.numeric(sub("mb", "0", Temp_C))
  ) %>%
  ggplot(aes(x = Temp_C, y = pct)) +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML",
                                 family = mgcv::tw(link = "log")),
              color = "black") +
  geom_point(
    aes(fill = Chamber),
    color = "black",
    shape = 21,
    stroke = 0.5,
    size = 2
  ) +
  chamber_scale +
  theme_hm +
  ylab(expression(atop("Ebullitive fluxes", "(% of all fluxes)"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4))


eb_magnitude <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12) %>%
  group_by(Chamber) %>%
  filter(
    !is.na(Ebullition_yn),
    Ebullition_yn
  ) %>%
  mutate(
    Temp_C = sub(
      " ºC", "",
      str_extract(
        as.character(Chamber),
        "-*[0-9]+\\.*[0-9]* ºC"
      )
    ),
    Temp_C = as.numeric(sub("mb", "0", Temp_C))
  ) %>%
  ggplot(aes(x = Temp_C, y = CH4_umol_m2_h)) +
  geom_point(
    aes(color = Chamber),
    shape = 21,
    stroke = 0.5,
    size = 2
  ) +
  scale_color_manual(
    limits = chamber_labels,
    values = color.gradient,
    name = "Chamber",
    drop = FALSE
  ) +
  theme_hm +
  ylab(expression(atop(
    "Ebullition magnitude",
    "(µmol" ~ m^-2 ~ h^-1 * ")"
  ))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4))

eb_annual <- top_pct %>%
  filter(name == "Ebullition events") %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(method = "gam",
              formula = y ~ s(x, k = 3),
              method.args = list(method = "REML",
                                 family = mgcv::tw(link = "log")),
              color = "black") +
  geom_point(
    aes(fill = Chamber),
    color = "black",
    shape = 21,
    stroke = 0.5,
    size = 2
  ) +
  chamber_scale +
  theme_hm +
  ylab(expression(atop(
    "Ebullitive emissions",
    "(% of growing season emissions)"
  ))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4))

comb <- patchwork::wrap_plots(
  a = eb_number,
  b = eb_magnitude + theme(legend.position = "none"),
  c = eb_annual+ theme(legend.position = "none"),
  nrow = 1,
  ncol = 3,
  #guides = "collect",
  axis_titles = "collect"
)
comb

ggsave(
  comb,
  filename = "../03_Figures/SI/SI - Ebullition.jpg",
  dpi = 300, width = 6, height = 3.5
)
```

## Eb contribution to extremes

```{r fig_3}
# Only use days that have data from all 12 chambers
hm_df <- ch4 %>%
  filter(!is.na(CH4_umol_m2_h)) %>%
  group_by(date) %>%
  filter(length(unique(Chamber)) == 12)

# Calculate the contribution of extreme fluxes
hot_moments_eb <- hot_moments_data %>%
  filter(!is.na(Ebullition_yn) & Ebullition_yn) %>%
  # Calculate sum of fluxes that exceed each threshold
  # and convert to growing season total (mol/m2/y from umol/m2/h)
  summarize(
    hot_sum_998 = sum(CH4_umol_m2_h[hot_moment_quant_998 == "hot"], na.rm = T) /
      1000 / 1000 * (n_days * 24),
    hot_sum_99 = sum(CH4_umol_m2_h[hot_moment_quant_99 == "hot"], na.rm = T) /
      1000 / 1000 * (n_days * 24),
    hot_sum_95 = sum(CH4_umol_m2_h[hot_moment_quant_95 == "hot"], na.rm = T) /
      1000 / 1000 * (n_days * 24),
    min = min(DateTime_EST),
    max = max(DateTime_EST)
  ) %>%
  mutate( # Add temperature treatment
    Temp_C = sub(" ºC", "", str_extract(as.character(Chamber), "-*[0-9]+\\.*[0-9]* ºC")),
    Temp_C = as.numeric(sub("mb", "0", Temp_C))
  )

# Calculate summary stats about the contribution of extreme events
top_pct_eb <- hot_moments_eb %>%
  left_join(hot_moments %>% select(Chamber, annual_flux, n)) %>%
  mutate(
    hm_pct_998 = hot_sum_998 / n / annual_flux * 100,
    hm_pct_99 = hot_sum_99 / n / annual_flux * 100,
    hm_pct_95 = hot_sum_95 / n / annual_flux * 100
  ) %>%
  select(Temp_C, Chamber, hm_pct_998, hm_pct_99, hm_pct_95) %>%
  pivot_longer(c(hm_pct_95, hm_pct_99, hm_pct_998)) %>%
  mutate(
    name = factor(
      name,
      levels = c("hm_pct_95", "hm_pct_99", "hm_pct_998"),
      labels = c(
        "Top 5% of measurements",
        "Top 1% of measurements",
        "Top 0.2% of measurements"
      )
    )
  )

top_pct_comb <- bind_rows(
  total = top_pct %>% filter(!name == "Ebullition events"),
  ebullition = top_pct_eb,
  .id = "Type"
)

# Plot the contribution of extreme events to total emissions
top_pct_fig <- top_pct_comb %>%
  mutate(name = droplevels(name)) %>%
  ggplot(aes(x = Temp_C, y = value)) +
  geom_smooth(
    aes(lty = Type),
    method = "gam",
    formula = y ~ s(x, k = 3),
    method.args = list(method = "REML"),
    color = "black"
  ) +
  geom_point(aes(fill = Chamber, shape = Type),
    color = "black",
    stroke = 0.5, size = 2
  ) +
  scale_shape_manual(values = c(22,21))+
  scale_linetype_manual(values = c("dotted","solid"))+
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab(expression(atop("Percent of growing season", "emissions"))) +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name, scales = "free_y")

top_pct_fig

# Save
ggsave(
  top_pct_fig,
  filename = "../03_Figures/Manuscript/Figure 3??? - Top percent contrib eb.jpg",
  dpi = 300, width = 6, height = 3.5
)

top_pct_fig <- top_pct_comb %>%
  mutate(name = droplevels(name)) %>%
  select(Temp_C, value, name, Chamber, Type) %>%
  pivot_wider(values_from = value, names_from = Type) %>%
  mutate(pct = ebullition/total*100) %>%
  ggplot(aes(x = Temp_C, y = pct)) +
  geom_smooth(
    method = "gam",
    formula = y ~ s(x, k = 3),
    method.args = list(method = "REML"),
    color = "black"
  ) +
  geom_point(aes(fill = Chamber),
    color = "black", shape = 21,
    stroke = 0.5, size = 2
  ) +
  scale_shape_manual(values = c(22,21))+
  scale_linetype_manual(values = c("dotted","solid"))+
  scale_fill_manual(values = color.gradient, name = "Chamber") +
  theme_hm +
  ylab("Percent ebullitive") +
  xlab("Temperature treatment (ºC above ambient)") +
  guides(fill = guide_legend(ncol = 4)) +
  facet_wrap(~name)+
  coord_cartesian(ylim = c(0,100))

top_pct_fig 

# Save
ggsave(
  top_pct_fig,
  filename = "../03_Figures/Manuscript/Figure 3??? - Top percent contrib eb2.jpg",
  dpi = 300, width = 6, height = 3.5
)
```

When does ebullition happen

```{r}
when <- ch4 %>%
  ggplot(aes(x = Depth_cm, y = CH4_umol_m2_h, color = Ebullition_yn))+
  geom_vline(aes(xintercept = 0)) +
  geom_point(alpha = 0.2, color = "grey50")+
  geom_point(data = . %>% filter(!is.na(Ebullition_yn) & Ebullition_yn),
             color = "red") +
  facet_wrap(~Chamber, scales = "free_y")+
  theme_hm

ggsave(
  when,
  filename = "../03_Figures/Manuscript/Figure ??? - Depth_eb.jpg",
  dpi = 300, width = 6, height = 4
)
```
